---
title: "Wholesale Customers Data Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document presents an analysis of the Wholesale Customers dataset, focusing on data transformation, descriptive statistics, and relationships between variables.

## Data Loading and Preparation

```{r libraries}
# Import necessary libraries
library(tidyverse)
library(tidyr)
library(tidyselect)
library(ggplot2)
library(data.table)
library(readxl)
library(readr)
library(psych)
library(writexl)
```

```{r data_loading}
# Reading the dataset
wholesale <- read_excel("Wholesale_customers_data.xls")
View(wholesale)

# Checking dataset structure
names(wholesale)
dim(wholesale)  # 440 rows, 8 columns
str(wholesale)
glimpse(wholesale)
```

## A. Data Transformation and Anomaly Handling

### Data Distribution Analysis

```{r histograms}
# Function to plot histograms for numeric columns
Histogram <- function(wholesale) {
  numeric_cols <- names(wholesale)[sapply(wholesale, is.numeric)] 
  
  par(mfrow = c(2, 2)) 
  
  for (col in numeric_cols) {
    hist(wholesale[[col]], main = paste("Histogram of", col), 
         xlab = col, col = "skyblue", border = "black")
  }
  
  par(mfrow = c(1, 1))  
}

# Calling the function
Histogram(wholesale)
```

### Missing Values Analysis

```{r missing_values}
# Checking for missing values
missing_values <- sapply(wholesale, function(x) sum(is.na(x)))
missing_values 
```

### Outlier Detection

```{r outliers}
# Function to detect outliers using boxplots
outlier_boxplot <- function(wholesale) {
  numeric_cols <- names(wholesale)[sapply(wholesale, is.numeric)] 
  
  par(mfrow = c(2, 2)) 
  
  for (col in numeric_cols) {
    boxplot(wholesale[[col]], main = paste("Boxplot of", col), 
         xlab = col, col = "skyblue", border = "black")
  }
  
  par(mfrow = c(1, 1))  
}

# Calling the function
outlier_boxplot(wholesale)
```

### Outlier Removal

```{r remove_outliers}
# Function to remove outliers
remove_outliers <- function(wholesale, cols) {
  for (col in cols) {
    q1 <- quantile(wholesale[[col]], 0.25, na.rm = TRUE)
    q3 <- quantile(wholesale[[col]], 0.75, na.rm = TRUE)
    IQR <- q3 - q1
    lower_bound <- q1 - 1.5 * IQR
    upper_bound <- q3 + 1.5 * IQR
    wholesale <- wholesale[wholesale[[col]] >= lower_bound & wholesale[[col]] <= upper_bound, ]
  }
  return(wholesale)
}

# Apply to numeric columns
numeric_cols <- names(wholesale)[sapply(wholesale, is.numeric)]
wholesale_cleaned <- remove_outliers(wholesale, numeric_cols)

# Verify outlier removal
outlier_boxplot(wholesale_cleaned)
```

## B. Save Transformed Dataset

```{r save_data}
# Save as R Dataframe
save(wholesale_cleaned, file = "Q12.RData")

# Save as Excel File
write_xlsx(wholesale_cleaned, "Q12.xlsx")
```

## C. Descriptive Statistics

```{r descriptive_stats}
# Select four variables for analysis
selected_vars <- wholesale_cleaned %>% 
  select(Fresh, Milk, Grocery, Detergents_Paper)

# Detailed descriptive statistics
describe(selected_vars)

# Summary statistics
summary(selected_vars)
```

## D. Relationship Between Continuous Variables

```{r correlation}
# Correlation test between Milk and Grocery
cor_test <- cor.test(wholesale_cleaned$Milk, wholesale_cleaned$Grocery)
print(cor_test)

# Visualization
ggplot(wholesale_cleaned, aes(x = Milk, y = Grocery)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", color = "red") +
  ggtitle("Relationship between Milk and Grocery")
```

## E. Relationship Between Categorical and Continuous Variables

```{r categorical_analysis}
# Convert Region and Channel to factors
wholesale_cleaned$Channel <- as.factor(wholesale_cleaned$Channel)
wholesale_cleaned$Region <- as.factor(wholesale_cleaned$Region)
``` 